HUME: Human UCCA-based Machine-Translation Evaluation

### 2016-03-09 paper/main.tex
aaa

  Semantic evaluation of Machine Translation (MT) systems has become
  a necessity, given recent interest in semantics-aware MT systems.
  We present a novel human semantic evaluation measure for MT, Human
  UCCA-based MT evaluation (HUME), building on the UCCA semantic representation scheme.
  Making progress over previous work, the proposed measure takes into account
  a wider range of semantic phenomena and does not rely on semantic annotation
  of the MT output, whose interpretation is often unclear. 
  We experiment with four language pairs, demonstrating the broad applicability of
  the measure, as well as its reliability, reflected by the inter-annotator agreement
  rates.

@@

Human judgement is the central, and perhaps ultimate, criterion for estimating the
quality of an MT system.
Nevertheless, common measures for MT evaluation, such as adequacy and frequency judgements
or the relative ranking of several possible translations, are problematic in two ways.
First, as the quality of translation is determined by multiple factors, it is difficult
to assign a single number that quantifies the quality of the entire sentence, as is reflected
by the diminishing inter-annotator agreement rates of human ranking measures
as a function of the sentence length \cite{Bojar:2011}.
Second, a global quality score does not localize the errors in the translation output,
and hence cannot inform measures to overcome these errors.
This problem can be addressed by using scores that decompose over
the sub-parts of the sentence. 
However, common decompositions into n-grams or even individual words, as employed by the
popular BLEU \cite{Papineni:2002} and METEOR measures \cite{Banerjee:2005},
are most suitable for evaluating shallow MT approaches and are
less instructive for more structurally-aware methods, which are at
the forefront of MT research today.

These difficulties have motivated the definition of a human measure that decomposes over
the semantic structure of the sentence, rather than its n-grams. The most notable attempt of
this brand is the HMEANT
measure \cite{LoWu2010,lo2011structured,LoWu2011b,LoWu2012}, which quantifies the
similarity the verbal argument structures of the translation output and reference.
While appealing in addressing the two major shortcomings
detailed above, HMEANT is unable to capture semantic differences which do not show up in
the verb-argument structure, including several very frequent phenomena, such as
inter-clause linkage, nominalizations and copula clauses (see \secref{sec:hmeant_comp}).
Another shortcoming of MEANT lies with its reliance on semantically annotating MT output,
which is often unintelligible. Finally, measures based on a comparison to a reference translation
is necessarily incomplete, given the huge number of reasonable translations
a sentence may have \cite{dryer_marcu:12}.

In this work we propose the Human UCCA-based MT Evaluation ({\it HUME}) metric,
a human evaluation measure that decomposes over the sentence's semantic units.
The units are defined according to the 
UCCA scheme \cite{abend2013universal}, an appealing candidate for semantic analysis,
due to its cross-linguistic applicability, support for rapid annotation, and coverage
of many fundamental semantic phenomena, such as verbal, nominal and adjectival
argument structures and the inter-relations between them.

@@

Semantic evaluation of machine translation has typically been done at the sentence level and
we propose an  approach which breaks down the evaluation into basic semantic units, making evaluation 
simpler and more consistent. 
Our  assumption is that the semantic structure on the source sentence should be 
retained in the translation, and if it is not, then some essential part of the meaning 
%will be 
is
lost. 
The semantic framework that we base our evaluations on is called 
Universal Conceptual Cognitive Annotation \shortcite{abend-rappoport:2013:ACL2013}.  
UCCA has been developed using linguistic theories about 
what types of components and structures are universal across many different languages.

>>>>

  Semantic evaluation of Machine Translation (MT) systems, informative as for
  semantic discrepencies between the translation and the source text,
  has become a necessity, given recent interest in semantics-aware MT systems.
  We present a novel human semantic evaluation measure for MT, Human
  UCCA-based MT evaluation (HUME), building on the UCCA semantic representation scheme.
  Making progress over previous work, the proposed measure takes into account
  a wider range of semantic phenomena and does not rely on semantic annotation
  of the MT output, whose interpretation is often unclear. 
  We experiment with four language pairs, demonstrating the broad applicability of
  the measure, as well as its reliability, reflected by the inter-annotator agreement
  scores.

@@

Human judgement is the central, and perhaps ultimate, criterion for estimating the
quality of an MT system.
Nevertheless, common measures for MT evaluation, such as adequacy and frequency judgements
or the relative ranking of several possible translations, are problematic in two ways.
First, as the quality of translation is determined by multiple factors, it is difficult
to assign a single number that quantifies the quality of the entire sentence, as is reflected
by the diminishing inter-annotator agreement rates of human ranking measures
as a function of the sentence length \cite{Bojar:2011}.
Second, a sentence-level quality score does not indicate what parts of the sentence
are erroneously translated in the output, and hence cannot inform developers in overcoming these errors.
This problem can be addressed by using scores that decompose over the sub-parts of the sentence. 
However, common decompositions into n-grams or even individual words, as employed by the
popular BLEU \cite{Papineni:2002} and METEOR measures \cite{Banerjee:2005},
are most suitable for evaluating shallow MT approaches and are
less instructive for more structurally-aware methods, which are at
the forefront of MT research today.

These difficulties have motivated the definition of a human measure that decomposes over
the semantic structure of the sentence, rather than its n-grams. The most notable attempt of
this type is the MEANT measure, and its human variant HMEANT
\cite{lo2010evaluating,lo2011structured,Lo2011meant,lo2013meant}, which quantifies the
similarity between the output translation and a reference translation in terms of the similarity
between their verbal argument structures
While the measure addresses the two major shortcomings
detailed above, it is unable to capture semantic differences which do not show up in
the verb-argument structure, including several very frequent phenomena, such as
inter-clause linkage, nominalizations and copula clauses (see \secref{sec:hmeant_comp}).
Another shortcoming of MEANT is its reliance on semantically annotating MT output,
which is often unintelligible. Finally, measures based on a comparison to a reference translation
are necessarily incomplete, as they compare to a small number of references
out of the the huge number of possible translations a sentence may have \cite{dryer_marcu:12}.

In this work we propose the Human UCCA-based MT Evaluation ({\it HUME}) metric,
a human evaluation measure that decomposes over the sentence's semantic units.
Semantic units are defined according to the 
UCCA scheme \cite{abend2013universal}, an appealing candidate for semantic analysis,
due to its cross-linguistic applicability, support for rapid annotation, and coverage
of many fundamental semantic phenomena, such as verbal, nominal and adjectival
argument structures and the inter-relations between them.

@@

Semantic evaluation of machine translation has typically been done at the sentence level and
we propose an  approach which breaks down the evaluation into basic semantic units, making evaluation 
simpler and more consistent. 
Our  assumption is that the semantic structure on the source sentence should be 
retained in the translation, and if it is not, then some essential part of the meaning 
%will be 
is
lost. 
The semantic framework that we base our evaluations on is called 
Universal Conceptual Cognitive Annotation \shortcite{abend2013universal}.  
UCCA has been developed using linguistic theories about 
what types of components and structures are universal across many different languages.
